### 2013-04-20 docs/paper_lncs.tex
Add lncs formatted version

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{float}
% \usepackage{caption}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{url}
\usepackage{cite}
\urldef{\mailsa}\path|{rcoh,  georgiou, prazavi}@mit.edu|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{A Reimplementation of Python’s dict using Simple Tabulation Hashing and Linear Probing}

% a short form should be given in case it is too long for the running head
\titlerunning{A Reimplementation of Python’s dict}

% the name(s) of the author(s) follow(s) next
\author{Russell Cohen\and Thomas Georgiou\and Pedram Razavi}
%
\authorrunning{A Reimplementation of Python’s dict}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Massachusetts Institute of Technology\\
77 Massachusetts Avenue  Cambridge, MA 02139, USA\\
\mailsa\\
\url{http://www.mit.edu}}

\maketitle

\begin{abstract}
Here we implement simple tabulation hashing with linear probing in Python
dictionary based on the recent theoretical result of P\v{a}tra\c{s}cu et al. Although
we achieved significant performance wins on synthetic benchmarks, on real
benchmarks our wins were only on the order of 4\% due to interpreter overhead.
Linear probing with sample tabulation hashing does improve Python performance
and should be looked at as a replacement scheme.  As a whole, the
extra memory bandwidth
and latency required to compute the simple tabulation hash negates some of the
performance improvement, but if they were able to be further mitigated, the new scheme
should provide more than a 4\% improvement, making it even more worthwhile.
\keywords{Hashing, Simple Tabulation}
\end{abstract}

\section{Introduction}
In this project we aim to analyze and improve the performance of Python's
dictionary or dict by changing its underlying implementation.  Python \cite{python} is an extremely popular general purpose computing language for several reasons:
\begin{enumerate}
 \item Clean, elegant near-pseudo code syntax leads to clear code 
 \item Widely distributed and available
 \item Large quantity of external libraries for specific features 
\end{enumerate}

Recently, with the advent of packages like numPy, sciPy and pyPy Python has began to also be used (with questionable merit) as a performance computing language due to its ease of use.

However, because Python is an interpreted language it is difficult to operate on a similar performance level to compiled languages due to the sheer overhead of the interpreter.

In order to attempt to improve the performance of the Python language as a whole we plan to attempt to improve the performance of Python dictionaries.

\subsection{Motivation}
Python heavily utilizes dictionary data structure for several of its internal language features. This extensive internal use of dictionaries means that a running Python program has many dictionaries active simultaneously, even in cases where the user's program code does not explicitly use a dictionary.

Take for example, the following code which we ran in the interpreter:

\begin{verbatim}
>> i = 2 
>> b = i * 2
>> b
10
\end{verbatim}

One of the features that we added to the Python source was a conditional flag
that allowed us to view dictionary statistics (collisions, probes, accesses,
etc.) for a given invocation of the Python binary.  With that modification we
were able to see that this code used 92 lookups in string dictionaries and 24
lookups in general purpose dictionaries for a total of 116 dictionary hits. 116
hits in a 3 line piece of code which never uses dictionaries explicitly is quite
significant.  We explain this extensive dictionary usage below.

For instance, a dictionary can be instantiated to pass keyword arguments to a
function. This means a dictionary could potentially be created and destroyed on
every function call.  Some other examples of internal usage of dictionaries in
Python are as follows:

\begin{itemize}
\item Class method lookup
\item Instance attribute lookup and global variables lookup
\item Python built-ins module
\item Membership testing
\end{itemize}

Later in this paper, we compare the performance of our new dictionary
implementation for some of these special cases. The aforementioned applications
of the dictionaries further highlights the need for the fast instantiation and
deletion of dictionaries so that less memory is utilized once running a program.
Aside from the issue of memory usage, to further enhance the runtime of a
program we need fast key/value storage and retrievals which is the main focus of
this paper. Therefore it becomes clear that a better dictionary structure will
have significant effects on total Python performance.

\section{Current Python Implementation}

In this section we give a brief overview of the key implementation details of
dictionaries in CPython 2.7.3.  We should note that Python's dictionary supports
several different data types as keys. However as we noted earlier, the
dictionaries underlying class instances and modules have only strings as keys.
Therefore optimizing a dictionary which have string-only keys is crucial to the
runtime of any Python program. CPython accommodates for this optimization by
changing its dictionary lookup method as necessary. First when a dictionary is
instantiated, CPython uses a string-only lookup function until a search for
non-string is requested, in this case CPython falls back to a more general
lookup method and uses this general function onward. There are two main
optimizations in the string-only lookup function.  First, since the string
comparisons do not raise any exceptions by design, some of the error checking
logic can been removed. Second, the string-only lookup function can skip the
rich set of comparisons ( $<=$, $>=$, $<$, $>$, $==$, $!=$) which arbitrary data
types can provide since string type does not have these special cases.  As it
can be seen, string-only keys can dominate the runtime of a dictionary and
CPython uses some measures to optimize for this common case. Because of this
importance we predominantly focus on the study of new implementations of
hashes for the string type for the rest of the paper.

The current CPython dictionary implementation uses an iterative XORed polynomial
terms for calculating the hash of the strings. In this implementation the hash
value for sequential strings differ only in the least significant bits, for
instance: hash("6.851") = 7824665118634166871 and hash("6.852") =
7824665118634166868. This behavior gives good results for sequential strings
which is a common case. When the table size is $2^i$ taking the least
significant $i$ bits as the index can populate the table without a lot of
collisions. This approach gives better than random results in this case and is
also simple and fast to compute.

On the other hand this hashing strategy has the tendency to fill contiguous
slots of the hash table. CPython tries to solve this behavior by utilizing a
good collision resolution strategy. The collision resolution strategy currently
used is open addressing with a custom non-linear probing scheme. In this scheme
CPython uses quadratic probing based on this function: $j = (5j + 1) \bmod 2^i$
where $0<j<2^i-1$. This recurrence alone is not enough for a better behavior
since again it behaves the same as linear-probing for some keys since it scans
the table entries in a fixed order. To overcome this issue CPython also uses the
most significant bits of the hash value with the following code snippet:

\begin{verbatim}
slot = hash;
perturb = hash;
while (<slot is full> && <item in slot does not equal the key>) {
       slot = (5*slot) + 1 + perturb;
       perturb >>= 5;
}
\end{verbatim}

Based on this new method the probe sequence will depend on nearly all the bits
of a hash value. We should note that the value of  perturb shift (currently 5)
is crucial for a good probing behavior.  It should be small enough so that
all the significant bits have an effect on the probe sequence but at the same
time it should be large enough so that in really bad cases the high-order hash
bits can affect the earlier iterations.

CPython initializes its table with 8 slots. The resizing strategy is based on a
load factor invariant. Whenever there are $n$ keys in the dictionary, there must
be at least $3n/2$ slots in the table. This load factor is to keep the table
sparse and put a bound on the number of collisions that can happen. CPython
resizes the size of table whenever the load factor invariant gets violated. It
quadruples the size when there are less than 50,000 keys in the table and
doubles the size otherwise.

Overall, the design of CPython dictionary is simple and is optimized for more
common cases such as sequential strings. Since it utilizes quadratic probing
with perturbation, even if a data type does not provide a good hash function,
the table still gets populated with with short chain lengths.  On the other hand
there are some behaviors which can be improved. For instance there is a bad
cache locality because for probing a slot the table might need to jump to the
other slots which are not immediate neighbors and not necessarily in cache.

\section{Improvements}

\subsection{Theoretical Results}
Recently, P\v{a}tra\c{s}cu et al. \cite{mihai} showed that simple tabulation hashing provides unexpectedly strong guarantees and is constant competitive with a truly random hash function for hashing with linear probing. In their paper, they showed that for simple tabulation hashing with linear probing if the table size is $m = (1 + \epsilon)n$, then the expected time per operation is $O(1/\epsilon^2)$. This new guarantee asymptotically matches the
bound for a truly random function. They also showed some experimental evaluations which proved that simple tabulation on fixed input length can be fast and robust when used with linear probing and cuckoo hashing.

Linear probing is appealing because of its excellent cache behavior in
comparison to the probing scheme currently utilized in Python.  Consider the
Intel Core 2 Duo processor with a 64 byte cache line.  Each Python dictionary
entry is 12 bytes so when Python probes a dictionary slot, it pulls in at least
4 more dictionary entries with it.  However, when the current probing scheme
probes another slot, it is not in the current cache line so it has to pull
another cache line in from memory.  With linear probing, that next slot is most likely in the current cache line.

P\v{a}tra\c{s}cu et al. \cite{mihai} describe simple tabulation hashing for fixed key widths.  For implementing this new hashing scheme, first we needed to figure out a variant of simple tabulation hashing for variable length strings that does not require a table for every possible index (we could potentially cycle through a fixed set of tables).  Performing table lookups is potentially an expensive operation with lots of cache hits mitigating the better cache behavior of linear probing so we will have to be careful about how to implement this. In the following section we explain the implementation details of our approach to these problems.

\subsection{Implementation Details}
We modified the source code of CPython 2.7.3 to use simple tabulation hashing
with linear probing.  We use a fixed number of tabulation tables and cycle between them
for variable length strings.

We prepopulated random tables in memory with random bits from random.org, which
sources randomness from atmospheric noise in order to remove the reliance on
pseudo-random number generators.

      Following is an excerpt from our hashing code.  We use bit tricks in
      order to prevent costly modulo instructions.  For each byte in the
      string, we construct an index by adding the byte value and a table offset
      together.  Then we lookup the value in a large table of random 64 bit
      longs and then XOR it with the previous result in the chain of XORs.

       \begin{verbatim}
    while (--len >= 0) {
        register long index = (*p++) | ((len&TABLE_MASK)<<8);
        x = x ^ RAND_TABLE_NAME[index];
    }
       \end{verbatim}

We also experimented with indexing using every pair of bytes in the string,
treating them as 16bit integers and
using $2^{16}$ tabulation table entries.

To implement linear probing, we modified Python's dictionary code to start at
the table entry defined by the hash value and then search sequentially until it
finds the item it is looking for or an empty cell.
\subsubsection{Optimizations}

After implementation and profiling, we found that dictionary table lookups had
over twice as many level 1 cache misses with tabulation hashing than without.
Hypothesizing that the tabulation tables were evicting other, more relevant
data from the level 1 cache.  To fix this, we inserted x86 memory prefetch
instructions into our hash code to tell the CPU's caching system that it should
deprioritize storing the tabulation tables in cache.  We used GCC's
\begin{tt}\_\_builtin\_prefetch(void \*addr, char rw, char locality)\end{tt}
to generate these instructions.  However, this only yielded about a 1\% performance improvement,
negligible compared to the performance improvement of different hash schemes.

\section{Results}

In order to analyze the result of this new dictionary we measured three main
metrics. First, we compared the performance of the simple tabulation hashing
with Python’s current hashing scheme on random and sequential strings. Second, we measured the number of probes and collisions and overall the quality of our dictionary on structured and unstructured inputs. Finally, in this section  we show some of the total time spent in dictionary operations by using a set of general Python benchmarks and some real-world programs.
\subsection{Table Number Tuning}

For simple tabulation hashing with variable length strings, we decided to use a
fixed number of tables $N$ and round robin between them for every byte.  We
found the optimal value of $N$ to be 8 experimentally.  We limited our choices
to powers of 2 so that costly modulo operations could be avoided by using bit
masks Fig.~\ref{fig:tables} shows the performance of different numbers of
tabulation tables with both 1 and 2 byte indices.  Eight hash tables indexing
every byte separately resulted in the best performance.    We believe this is due
to the low entropy caused by fewer tables and too much memory bandwidth used by a higher
number of tables.  The optimal result may also be an artifact of the fact that
we were using on the order of magnitude of thousands to millions of dictionary
keys.  Our optimal result resulted using 256 64
bit table entries in each of 8 tables using 16KB of memory.
 \begin{figure}
   \centering
   \includegraphics[width=4.1in]{tables.pdf}
   \caption{Eight tabulation tables with 1 byte indices provides the best performance for a
   dictionary with 6 million keys}
   \label{fig:tables}
 \end{figure}

\subsection{Synthetic Benchmarks}

Our primary benchmark, hash-table-shootout, is a synthetic one that called the CPython dictionary
code directly from C so that all the interpreter overhead was bypassed.  We
measured performance for both sequential strings, generated by converting
sequential integers into their string representation, and for random strings,
generated by the string representation of random integers.  For each of the
benchmarks, we inserted a given number of keys into the dictionary and measured
the runtime.

Fig.~\ref{fig:randstring} shows the performance for random strings.  Both
linear probing and simple tabulation result in improved performance over the
current CPython dictionary implementation but combining both techniques results
in over a 20\% improvement.  Fig.~\ref{fig:seqstring} shows very similar
results for sequential strings.

% \noindent%
% \begin{minipage}{\linewidth}
% \makebox[\linewidth]{%
%   \includegraphics[width=4.1in]{randstring.pdf}}
%   \captionof{figure}{Random string performance}
%   \label{fig:randstring}
% \end{minipage}

% \noindent%
% \begin{minipage}{\linewidth}
% \makebox[\linewidth]{%
%   \includegraphics[width=4.1in]{seqstring.pdf}}
%   \captionof{figure}{Sequential string performance}
%   \label{fig:seqstring}
% \end{minipage}

\begin{figure}[H]
  \centering
  \includegraphics[width=4.1in]{randstring.pdf}
  \caption{Random string performance}
  \label{fig:randstring}
\end{figure}
\begin{figure}[H]
  \centering
  \includegraphics[width=4.1in]{seqstring.pdf}
  \caption{Sequential string performance}
  \label{fig:seqstring}
\end{figure}

Fig.~\ref{fig:chainlength} shows the average chain length, or dictionary
cell probes per lookup for a run of the benchmark on sequential strings.
Simple tabulation hashing with CPython's current collision resolution strategy
results in the fewest number of probes.  However, despite having the fewest
number of probes, its real relative performance is much worse as seen in Fig.~\ref{fig:seqstring}.  Linear probing performs faster than its average
chain length suggests, showing the impact of looking at entries in the same
cache line first rather than jumping around.

\begin{figure}
  \centering
  \includegraphics[width=4.1in]{chainlength.pdf}
  \caption{Simple tabulation hashing decreases the average chain length}
  \label{fig:chainlength}
\end{figure}

\subsection{Benchmarks}
Due to the heavy utilization of dictionaries in Python internals, one of the
valuable benchmarks is to see how the new dictionary performs with them them. We used
PyBench benchmark suite for this task. The results can be seen in Table
~\ref{pybenchtable}.  We saw significant improvements of around
20\% in performance of some of the internals such as the special and normal
class attribute lookups and string predicates. However, at the same time some
of the other internals such as CPython exception handler were running around
12\% slower compared to the current implementation. On average a 4.5\% improvement in runtime was measured for the whole benchmark suite.

\begin{table}
\centering
  \caption{Performance results on selected components of the PyBench suite.  We
    have omitted the least significant performance results in terms of percent
    change.
  }
\begin{tabular}{| l | c |}
  \hline
  \textbf{Test Name} & \textbf{Performance Difference} \\ \hline
  NormalClassAttribute &  19.20\% \\ \hline
  SpecialClassAttribute &  18.88\% \\ \hline
  NormalInstanceAttribute &  17.77\% \\ \hline
  StringPredicates &  15.25\% \\ \hline
  UnicodePredicates &  13.82\% \\ \hline
  SimpleListManipulation &  13.03\% \\ \hline
  CompareUnicode &  11.27\% \\ \hline
  SimpleComplexArithmetic &  10.83\% \\ \hline
  SpecialInstanceAttribute &  10.52\% \\ \hline
  TryFinally &  10.35\% \\ \hline
  CreateUnicodeWithConcat &  8.52\% \\ \hline
  StringSlicing &  6.52\% \\ \hline
  Small Performance Changes Omitted & ... \\ \hline
  UnicodeMappings &  -1.27\% \\ \hline
  ListSlicing &  -2.40\% \\ \hline
  ComplexPythonFunctionCalls &  -3.55\% \\ \hline
  TupleSlicing &  -3.72\% \\ \hline
  WithRaiseExcept &  -5.89\% \\ \hline
  TryRaiseExcept &  -11.99\% \\ \hline
  Totals & 4.54\% \\ \hline
\end{tabular}
\label{pybenchtable}
\end{table}

PyBench showed the performance difference in each of the individual CPython
internals however for a more realistic picture of the performance in real world
programs more extensive benchmarking of the new implementation was needed.
Especially since we wanted to know the frequency of the different CPython
internals in real-world applications and therefore the corresponding runtime
changes. The new implementation was tested on several standard real world
programs such as Tornado web server, Django web framework, html5lib and ten other real-world programs.

We were interested to calculate if the new implementation of dictionary can handle more requests in a Python web server compared to the current CPython implementation. We chose Tornado web server for this task since most of the code is written in pure Python \cite{wsgi}. An instance of Tornado server was run in a virtual machine with one assigned processor. The code of the application used is provided in the appendix.

Different request rates (up to 2000 requests per second) were tested on this
sample server and their corresponding response rates were calculated using
httperf, a HTTP performance measurement tool. Considering the variance of request rates on different iterations of the test, we could not find any significant improvement of the response rate of the new implementation. We found that on the average performance was similar however the new implementation showed a smaller standard deviation in the response rates.

This finding as well as those by other benchmarks showed that the new
implementation gives more consistent runtime on different iterations of a same
code compared to the current CPython implementation. This consistency is a
desired behavior. We suspect that this behavior is a result of the ``better'' randomness of hash values in simple tabulation hashing and also the smaller average chain lengths.

As mentioned before, another set of the benchmarks \cite{hashshoot} that we utilized was based on different standard real-world Python applications. We ran and timed the performance of the new implementation and CPython 2.7.3 side by side using the Python-Mirrors suite. The performance of more than ten applications were recoded as shown in the figures.

\begin{figure}
  \centering
  \includegraphics[width=12cm]{appsbench.pdf}
  \caption{Percentage of overall improvements in some real-world Python
  applications. Nearly none of the applications performed worse. Python Pickle
  and html5lib faster around 3\% to 4\%. }
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=12cm]{slowpickle.png}
  \caption{Python Pickle module benchmark. Minimum Iteration: changed from 0.705s to 0.681s: 1.04 times faster. Average iteration: changed from 0.710s to 0.686s: 1.04 times faster. Standard deviation of new implementation 2.0383 times smaller}
\end{figure}

% \begin{figure}[H]
%   \centering
%   \includegraphics[width=160mm]{lr-10.png}
%   \caption{}
% \end{figure}

% \begin{figure}[H]
%   \centering
%   \includegraphics[width=160mm]{lr-10.png}
%   \caption{}
% \end{figure}

In a couple of the programs such as pure Python Pickle (an object serialization
library) and html5lib (specification parser library) we calculated an
improvement of 3\% to 4\%. In other cases we did not calculate an speedup but
at the same time we did not find any real-world case which the new
implementation performs significantly slower. These results were consistent
with the previous results of PyBench that showed about 4.5\% improvement.

\section{Conclusion and future work}
Throughout the course of our work we came to the following conclusions:
\begin{enumerate}
\item Hardware specific performance issues dominate at small scales: When you
are attempting to work on a process that runs as fast as a hash table lookup or
insertion already does, there is little room for error.  Improving hash
performance with simple tabulation is not an easy task simply because any
change you make will probably end up making the hash function run slower.  That said:

\item Simple tabulation yields slightly faster and more consistent performance:
With a careful implementation simple tabulation can yield increased hash
performance on linear probing work loads. Simple tabulation definitely leads to
shorter chain lengths, however, those shorter chain lengths don’t always
justify the performance cost in all cases.

\item Memory latency presents serious but workable issues.  Doing a look up in the simple-tabulation table \emph{requires} a random memory access, exactly what we were trying to avoid. With careful memory management we can maintain good performance, but a solution requires careful thought about cache sizes and special purpose compiler directives.

\end{enumerate}

In terms of future work, we have 2 avenues of attack. First, we could use
different dictionaries for different purposes. Many of the Python dictionaries
are small, and their keysets never change. New built in functions are not added
during the runtime of the program and thus they could use perfect hashing.

Second, we could improve the performance bottlenecks illustrated by benchmarks.
These mainly consist of the extra time and memory bandwith and latency cost for
computing simple tabulation hashing.  If we optimize the hashing further
through more profiling, memory layout, and instruction order optimizing, we may
be able to achieve more real world gains.  We could also look into applying a
modifying of the perturbation trick currently used in Python for quadratic
probing.

Given the strong performance of simple tabulation hashing and the only downside
being the cost of memory lookups, processor manufacturers should look into
adding random tables into hardware as a ROM.  We found that for what we found
optimal for Python, we only needed 16KB of randomness.  This would be a trivial
cost in ROM given that processors have MBs of higher level caches nowadays.

\bibliography{references}{}
\bibliographystyle{plain}
\end{document}

\appendix
\newpage
\section*{Appendix: Tornado Application}
Sample Tornado 2.2.1 application used for benchmark:
 \begin{verbatim}
import tornado.httpserver
import tornado.ioloop
import tornado.wsgi

def application(environ, start_response):
    status = '200 OK'
    output = 'Hello World!' 
    response_headers = [('Content-type', 'text/plain'),
        ('Content-Length', str(len(output)))]
    start_response(status, response_headers)
    return [output]
def main():
    container = tornado.wsgi.WSGIContainer(application)
    http_server = tornado.httpserver.HTTPServer(container)
    http_server.listen(8000)
    tornado.ioloop.IOLoop.instance().start()
if __name__ == "__main__":
    main()
 \end{verbatim}
\appendix

\end{document}